---
title: "2.1Convert"
author: "Timo Kelder"
date: "June 4, 2020"
output: github_document 
---

In this notebook, we import the netcdf files genereted in the first step of the workflow and convert them to R dataframes. 

## Import data and packages

```{r load-packages, include=FALSE}
dir='C:/Users/Timo/OneDrive - Loughborough University/GitHub/UNSEEN-open/Data'

library('ncdf4')
library('tidyverse')
library('plyr')
library(knitr)

```

### Load the data
This data is downloaded from CDS. see the Notebooks of the first step of the workflow. We load the netcdf of SEAS5 data, that contains February precipitation forecasts for all 25 members over 35 years (1981-2016) and and 5 leadtimes. 
```{r}
SEAS5_UK_weighted_masked_nc = nc_open(paste0(dir,'/SEAS5_UK_weighted_masked.nc')) # Open the netcdf file
SEAS5_UK_weighted_masked_array = ncvar_get(SEAS5_UK_weighted_masked_nc) # and get the values
dim(SEAS5_UK_weighted_masked_array) ## Show the dimensions of the netcdf file
```
The dimnames do not get loaded from the netcdf created with Xarray. Here I set the dimnames. The dimensions are:

- Member: 0-24
- Year: 1982-2016
- Leadtime: 2-6

```{r}
dimnames(SEAS5_UK_weighted_masked_array)
dimnames(SEAS5_UK_weighted_masked_array) = list(as.character(0:24),1982:2016,as.character(2:6))
names(dimnames(SEAS5_UK_weighted_masked_array)) <- c('Member', 'Year','Leadtime')
```

R works mostly in dataframes rather than arrays (e.g. ggplot and tidyverse). Therefore, we convert the array into a dataframe.

```{r}
SEAS5_UK_weighted_masked_df = adply(SEAS5_UK_weighted_masked_array, 1:3) ## Convert the array to a data frame. Split up data by all dimensions (1:3)  
```

Rename the variable (V1) into ' Precipitation' and change the class of the 'year' column into numeric. 

```{r}
SEAS5_UK_weighted_masked_df = SEAS5_UK_weighted_masked_df %>%
  dplyr::rename(Precipitation = V1)

head(SEAS5_UK_weighted_masked_df)
SEAS5_UK_weighted_masked_df$Year = as.numeric(levels(SEAS5_UK_weighted_masked_df$Year))[SEAS5_UK_weighted_masked_df$Year]
SEAS5_UK_weighted_masked_df$Member = as.numeric(levels(SEAS5_UK_weighted_masked_df$Member))[SEAS5_UK_weighted_masked_df$Member]
SEAS5_UK_weighted_masked_df$Leadtime = as.numeric(levels(SEAS5_UK_weighted_masked_df$Leadtime))[SEAS5_UK_weighted_masked_df$Leadtime]

head(SEAS5_UK_weighted_masked_df)
```

Check if the conversion was successful: select member = 0, year = 1981 and Leadtime = 2. 
```{r}
SEAS5_UK_weighted_masked_array['0','1982','2']
SEAS5_UK_weighted_masked_df[1,,]
```

And the EOBS observations:

```{r}
EOBS_UK_weighted_upscaled_nc = nc_open(paste0(dir,'/EOBS_UK_weighted_upscaled.nc'))

EOBS_UK_weighted_upscaled_array = ncvar_get(EOBS_UK_weighted_upscaled_nc) # and get the values
dim(EOBS_UK_weighted_upscaled_array) ## Show the dimensions of the netcdf file
dimnames(EOBS_UK_weighted_upscaled_array) = list(1950:2020) #set the dimnames as the years 1950-2020
names(dimnames(EOBS_UK_weighted_upscaled_array)) <- 'Year'

EOBS_UK_weighted_upscaled_df = data.frame(EOBS_UK_weighted_upscaled_array) ## Convert the array to a data frame.
EOBS_UK_weighted_upscaled_df = rownames_to_column(EOBS_UK_weighted_upscaled_df, var="Year")
EOBS_UK_weighted_upscaled_df$Year = as.numeric(EOBS_UK_weighted_upscaled_df$Year)
head(EOBS_UK_weighted_upscaled_df)

EOBS_UK_weighted_upscaled_df = EOBS_UK_weighted_upscaled_df %>%
  dplyr::rename(Precipitation = EOBS_UK_weighted_upscaled_array)
head(EOBS_UK_weighted_upscaled_df)


#check
EOBS_UK_weighted_upscaled_array['1959']
EOBS_UK_weighted_upscaled_df[10,2]

```

```{r}
write.csv(EOBS_UK_weighted_upscaled_df,'../../Data/EOBS_UK_weighted_upscaled_df.csv')
write.csv(SEAS5_UK_weighted_masked_df,'../../Data/SEAS5_UK_weighted_masked_df.csv')
```
